Ignition of Fuel Beds by Firebrands

### 2022-03-30 manuscript1.tex
Multi-heater introduction draft completed. It needs significant work but the structure is there.

    Studies evaluating the heat flux of firebrands and the critical heat flux for ignition have yielded further insights into the ignition process. Hakes et al.~\cite{Hakes2019a} found that, for a single cylindrical firebrand and piles of firebrands, peak heat flux values ranged between 20 and 60 \si{\kilo\watt\per\square\meter} with average heat fluxes between 12 and 25 \si{\kilo\watt\per\square\meter}. The mass of the firebrands or piles of firebrands had little effect on the peak heat flux but directly influenced the total energy released. Tao et al.~\cite{Tao2020} observed similar heat fluxes for various of natural and manufactured firebrands. Both Hakes et al.~\cite{Hakes2019a} and Tao et al.~\cite{Tao2020} observed that an increase in wind speed significantly increased the measured heat flux. Hernandez et al.~\cite{Hernandez2018} found that Monterey Pine (\textit{pinus radiata}) needles ignited under heat flux as low as 10 \si{\kilo\watt\per\square\meter} with ignition time decreasing proportionally to the inverse square of increasing heat flux. In similar tests but with different fuels, Rivera et al.~\cite{Rivera2020} observed that critical heat fluxes for ignition were highly dependent on fuel bed properties with the critical radiative heat flux increasing as the porosity decreased. Reported critical values ranged from 6.64 \si{\kilo\watt\per\square\meter} to 20.85\si{\kilo\watt\per\square\meter} for Monterey Pine needles with porosities of 0.09 and 0.01, respectively. It has been observed that a variety of firebrands are capable of producing heat fluxes well above the critical heat flux values long enough for ignition in some fuels. However, upon comparing these values to other studies, ignition is not guaranteed if the critical heat flux rate and duration are met. For example, experiments conducted by Manzello et al.~\cite{Manzello2008} used firebrands similar to Hakes et al.~\cite{Hakes2019a} and Tao et al.~\cite{Tao2020} with fuels similar to Hernandez et al.~\cite{Hernandez2018} and Rivera et al.~\cite{Rivera2020} (e.g., wooden disks on pine needles) but did not observe ignition under conditions that would be anticipated to produce ignition. It should be noted that the studies conducted by Hernandez et al. and Rivera et al. were conducted under quiescent conditions and those by Manzello et al. between 0.5\si{\meter\per\second} and 1.0\si{\meter\per\second}. Nevertheless, the reported values of firebrand heat flux at 0.5\si{\meter\per\second} and 1.2\si{\meter\per\second} conditions by Tao et al. suggest ignition is likely to occur for instances where no ignition was observed. Not observing ignition under conditions at the apparent intersection of these findings suggests that other factors may be as important as heat flux and duration of heating.

@@

\section*{Methodology}
   The time to ignition was measured for five different fuel bed conditions with varying surface temperatures of a resistance heater.  The time to ignition was the metric used to evaluate the ignition propensity. The experimental apparatus, as illustrated in Figure~\ref{fig:ignPropensity}, was designed to replicate both conduction and radiation that may occur when a firebrand lands on the fuel bed. The heater was held in place by a lever arm that, when lowered, positioned the heater at a fixed location for the duration of the test. The firebrand was represented by a 6.35\si{\milli\meter} diameter 51\si{\milli\meter} long cartridge heater capable of a 250\si{\watt} output. The heater was inserted 3\si{\milli\meter} into the bed (approximately half the diameter) in the porous media tests and on top of the plates for the other experiments. The temperature of the heater was continually recorded via a type-K thermocouple attached to the top of the heater. An important distinction between using the lever arm holder and a naturally occurring firebrand is that the location of the heater remained fixed and, for times greater than roughly 10\si{\second}, could lose contact with the fuel bed as material was lost because of pyrolysis.  Thus, for the longer ignition experiments the arrangement mimicked a firebrand with a gap between it and the fuel bed, instead of a firebrand that maintained consistent contact. The rationale in using the lever arm was to ensure that the heater was placed a consistent depth within the fuel bed since sensitivities of ignition to heat source penetration depth have been observed by Wang et al.~\cite{Wang2016}. The temperature of the heater was held to within $\pm$ 6\% of the set point using PID control implemented in LabVIEW. Power delivery to the heater was measured at a rate of 1kHz for all tests. 
   Admittedly, the temperature and heat transfer from an actual firebrand to a fuel bed may vary more than that of a controlled heater, nor does the heater have a piloted ignition source. Nonetheless, trends of ignition propensity are expected to be similar between the heater and firebrands since the heat transfer rates calculated in these experiments are in the range of 1\si{\kilo\watt\per\square\meter} to 21\si{\kilo\watt\per\square\meter} which are comparable to heat flux values reported by Hakes et al.~\cite{Hakes2019a} and Tao et al.~\cite{Tao2020} for combustion of glowing firebrands on an instrumented surface. The advantage of using a heater was that it allowed sensitivities of ignition to the fuel beds and controlling processes to more readily be identified because the boundary conditions were measured, controlled, and consistent.

@@

\section*{Results}
   The time required for flaming ignition to occur for the various fuel beds is shown in Figure~\ref{fig:ignTime} as a function of heater set point temperature. Four observations are noted. First, the ignition times generally occurred  within the first 10\si{\second}.  If ignition did not occur after 10\si{\second} then it would typically take between 100\si{\second} and 1000\si{\second} to ignite, if at all. Conditions where ignition did not occur are not included in Figure~\ref{fig:ignTime}. A histogram of ignition times and the probability density for each material are shown in Figure~\ref{fig:ignHist} to further quantify the distribution of ignition times. The probability density of the $L_{c}<$ 1\si{\milli\meter} fuel particles, Douglas-fir plates, and cardboard plates are normally distributed with centers at 2.3\si{\second}, 2.8\si{\second}, and 3.9\si{\second}. The $L_{c}<$ 1\si{\milli\meter} fuel particles have an outlier peak centered at 1000\si{\second}. The 4\si{\milli\meter} $<L_{c}<$ 6\si{\milli\meter} and 6\si{\milli\meter} $<L_{c}<$ 12\si{\milli\meter} fuel particles are bimodal with highest density peaks at 1.7\si{\second} and 113\si{\second} respectively. The secondary peaks occur at 113\si{\second} for the 4\si{\milli\meter} $<L_{c}<$ 6\si{\milli\meter} fuel particles and 2.1\si{\second} for the 6\si{\milli\meter} $<L_{c}<$ 12\si{\milli\meter} fuel particles. Second, the probability of ignition at extended times increased as the particle sizes increased. Specifically, the proportion of ignition events in where $t_{ign} <$ 10\si{\second} group were 90\%, 77\%, and 47\% for the particles $L_{c}<$ 1\si{\milli\meter}, 4\si{\milli\meter} $<L_{c}<$ 6\si{\milli\meter}, and 6\si{\milli\meter} $<L_{c}<$ 12\si{\milli\meter} particle sizes, respectively. The third observation is that ignition was not observed beyond 100s for either of the solid plate fuel bed materials. Trends in ignition times for the plates were most similar to those for beds with the smallest particles. Fourth, for the ignition events that occurred within the first 10\si{\second} there is no apparent relationship between time to ignition, temperature, particle size, and fuel bed type. Additionally, the long timescales of some ignition events suggest that smoldering initiates and then transitions to flaming combustion. Since the incidence of ignition at extended times increases as the particle size increases the potential for smoldering to flaming transition is attributed to thermal and physical properties of the fuel bed. The different sensitivities of ignition just described are attributed to differences in the bulk thermal properties, the interface between the heater and fuel bed, and the global equivalence ratio of the fuel bed, as explained later. 
        \begin{figure}[htp]
            \centering
            \begin{tikzpicture}
                \node[anchor=south west,inner sep=0] (image) at (0,0) {\includegraphics[width=\figureWidthSet]{Figures/ignition_comparison.png}};
                    \begin{scope}[x={(image.south east)},y={(image.north west)}]
                        \draw[dashed, thick,rounded corners] (0.12, 0.17) rectangle (0.33, 0.80);
                        \draw[dotted, thick,rounded corners] (0.46, 0.17) rectangle (0.68, 0.80);
                    \end{scope}
            \end{tikzpicture}
            % \includegraphics[width=\figureWidthSet]{Figures/ignition_comparison.png}
            \caption{Time to ignition and heater temperature at ignition for all fuel bed materials. 
            The dashed and dotted boxes emphasize the two general times-scales associated with ignition.}
            \label{fig:ignTime}
        \end{figure}

@@

\section*{Summary and Conclusions}
    Flaming ignition tests have been conducted for porous Douglas-fir beds, Douglas-fir plates, and cardboard plates. A cylindrical cartridge heater was used as a firebrand surrogate. Heater temperature and electrical power to the heater were collected throughout each test. The derived heat flux to the fuel bed was within the range reported in literature of heat fluxes delivered by firebrands.  The time to ignition and probablity of ignition were used to evaluate the ignition propensity for the various fuel beds and heater temperatures.  A simplified heat transfer, pyrolysis, and ignition delay model was developed and used to provide further insights into the physical processes associated with ignition. The specific conclusions from this work are as follows.
        \begin{enumerate}
            \item 
                Smaller particles ignite more readily in porous beds than larger particles when heat transfer from the heater is primarily through conduction. This was evident by higher ignition probabilities, in general, of the smaller particles for a fixed heater temperature. As particle sizes increase radiant heat transfer becomes more important and fuel beds with larger particles were more likely than smaller particles to ignite at extended times (\textgreater 100\si{\second}) due to the increased importance of radiant ignition. 
            \item
                Douglas-fir plates ignite at times where conduction is the dominant mode of heat transfer (\textless 10\si{\second}) due to the higher thermal conductivity of the solid plates. The ignition probability of plates was the most similar to the larger particle, in particular at lower heater temperatures, due to dispersed heating of the porous fuel bed through radiation and the increased thermal conductivity of the plates creating similar temperature profiles. The rise in ignition probability  over a smaller heater temperature range time with temperature results from more consistent contact between the heater and plate surface.
            \item 
                Heat flux delivered to the fuel bed, when compared to heater temperature, is more indicative of ignition likelihood and ignition time for porous fuel beds. Heat flux is a more significant predictor of ignition because it captures differences in heat transfer modes and particle contact that heater temperature values do not. While this finding is not new, what is novel is that the mixed mode of heating (conduction and radiation) has a significant impact on the flaming ignition of fuel beds.
            \item 
                Consideration of the transport characteristics of pyrolyzate gases near the high temperature source can be important for more fully predicting ignition propensity. A \textit{Da} of ignition, in relation to the measured heat flux and thermal diffusivity of the fuel beds, is a promising relationship for predicting ignition for the porous fuel beds.  
        \end{enumerate}
    Further work is needed to verify that the \textit{Da} may be used to predict ignition for solid surfaces and for porous fuel beds with varying chemical compositions.  If proven valid, the (\textit{Da}), measured/predicted heat fluxes, and fuel bed properties may be used to help predict ignition of fuel beds both in and out of the WUI, ultimately helping to increase the effectiveness of fire prevention and suppression efforts.

>>>>

    Studies evaluating the heat flux of firebrands and the critical heat flux for ignition have yielded further insights into the ignition process. Hakes et al.~\cite{Hakes2019a} found that, for a single cylindrical firebrand and piles of firebrands, peak heat flux values ranged between 20 and 60 \si{\kilo\watt\per\square\meter} with average heat fluxes between 12 and 25 \si{\kilo\watt\per\square\meter}. The mass of the firebrands or piles of firebrands had little effect on the peak heat flux but directly influenced the total energy released. Tao et al.~\cite{Tao2020} observed similar heat fluxes for various of natural and manufactured firebrands. Both Hakes et al.~\cite{Hakes2019a} and Tao et al.~\cite{Tao2020} observed that an increase in wind speed significantly increased the measured heat flux. Hernandez et al.~\cite{Hernandez2017} found that Monterey Pine (\textit{pinus radiata}) needles ignited under heat flux as low as 10 \si{\kilo\watt\per\square\meter} with ignition time decreasing proportionally to the inverse square of increasing heat flux. In similar tests but with different fuels, Rivera et al.~\cite{Rivera2020} observed that critical heat fluxes for ignition were highly dependent on fuel bed properties with the critical radiative heat flux increasing as the porosity decreased. Reported critical values ranged from 6.64 \si{\kilo\watt\per\square\meter} to 20.85\si{\kilo\watt\per\square\meter} for Monterey Pine needles with porosities of 0.09 and 0.01, respectively. It has been observed that a variety of firebrands are capable of producing heat fluxes well above the critical heat flux values long enough for ignition in some fuels. However, upon comparing these values to other studies, ignition is not guaranteed if the critical heat flux rate and duration are met. For example, experiments conducted by Manzello et al.~\cite{Manzello2008} used firebrands similar to Hakes et al.~\cite{Hakes2019a} and Tao et al.~\cite{Tao2020} with fuels similar to Hernandez et al.~\cite{Hernandez2017} and Rivera et al.~\cite{Rivera2020} (e.g., wooden disks on pine needles) but did not observe ignition under conditions that would be anticipated to produce ignition. It should be noted that the studies conducted by Hernandez et al. and Rivera et al. were conducted under quiescent conditions and those by Manzello et al. between 0.5\si{\meter\per\second} and 1.0\si{\meter\per\second}. Nevertheless, the reported values of firebrand heat flux at 0.5\si{\meter\per\second} and 1.2\si{\meter\per\second} conditions by Tao et al. suggest ignition is likely to occur for instances where no ignition was observed. Not observing ignition under conditions at the apparent intersection of these findings suggests that other factors may be as important as heat flux and duration of heating.

@@

\section{Methodology}
   The time to ignition was measured for five different fuel bed conditions with varying surface temperatures of a resistance heater.  The time to ignition was the metric used to evaluate the ignition propensity. The experimental apparatus, as illustrated in Figure~\ref{fig:ignPropensity}, was designed to replicate both conduction and radiation that may occur when a firebrand lands on the fuel bed. The heater was held in place by a lever arm that, when lowered, positioned the heater at a fixed location for the duration of the test. The firebrand was represented by a 6.35\si{\milli\meter} diameter 51\si{\milli\meter} long cartridge heater capable of a 250\si{\watt} output. The heater was inserted 3\si{\milli\meter} into the bed (approximately half the diameter) in the porous media tests and on top of the plates for the other experiments. The temperature of the heater was continually recorded via a type-K thermocouple attached to the top of the heater. An important distinction between using the lever arm holder and a naturally occurring firebrand is that the location of the heater remained fixed and, for times greater than roughly 10\si{\second}, could lose contact with the fuel bed as material was lost because of pyrolysis.  Thus, for the longer ignition experiments the arrangement mimicked a firebrand with a gap between it and the fuel bed, instead of a firebrand that maintained consistent contact. The rationale in using the lever arm was to ensure that the heater was placed a consistent depth within the fuel bed since sensitivities of ignition to heat source penetration depth have been observed by Wang et al.~\cite{Wang2016}. The temperature of the heater was held to within $\pm$ 6\% of the set point using PID control implemented in LabVIEW. Power delivery to the heater was measured at a rate of 1kHz for all tests. 
   Admittedly, the temperature and heat transfer from an actual firebrand to a fuel bed may vary more than that of a controlled heater, nor does the heater have a piloted ignition source. Nonetheless, trends of ignition propensity are expected to be similar between the heater and firebrands since the heat transfer rates calculated in these experiments are in the range of 1\si{\kilo\watt\per\square\meter} to 21\si{\kilo\watt\per\square\meter} which are comparable to heat flux values reported by Hakes et al.~\cite{Hakes2019a} and Tao et al.~\cite{Tao2020} for combustion of glowing firebrands on an instrumented surface. The advantage of using a heater was that it allowed sensitivities of ignition to the fuel beds and controlling processes to more readily be identified because the boundary conditions were measured, controlled, and consistent.

@@

\section{Results}
   The time required for flaming ignition to occur for the various fuel beds is shown in Figure~\ref{fig:ignTime} as a function of heater set point temperature. Four observations are noted. First, the ignition times generally occurred  within the first 10\si{\second}.  If ignition did not occur after 10\si{\second} then it would typically take between 100\si{\second} and 1000\si{\second} to ignite, if at all. Conditions where ignition did not occur are not included in Figure~\ref{fig:ignTime}. A histogram of ignition times and the probability density for each material are shown in Figure~\ref{fig:ignHist} to further quantify the distribution of ignition times. The probability density of the $L_{c}<$ 1\si{\milli\meter} fuel particles, Douglas-fir plates, and cardboard plates are normally distributed with centers at 2.3\si{\second}, 2.8\si{\second}, and 3.9\si{\second}. The $L_{c}<$ 1\si{\milli\meter} fuel particles have an outlier peak centered at 1000\si{\second}. The 4\si{\milli\meter} $<L_{c}<$ 6\si{\milli\meter} and 6\si{\milli\meter} $<L_{c}<$ 12\si{\milli\meter} fuel particles are bimodal with highest density peaks at 1.7\si{\second} and 113\si{\second} respectively. The secondary peaks occur at 113\si{\second} for the 4\si{\milli\meter} $<L_{c}<$ 6\si{\milli\meter} fuel particles and 2.1\si{\second} for the 6\si{\milli\meter} $<L_{c}<$ 12\si{\milli\meter} fuel particles. Second, the probability of ignition at extended times increased as the particle sizes increased. Specifically, the proportion of ignition events in where $t_{ign} <$ 10\si{\second} group were 90\%, 77\%, and 47\% for the particles $L_{c}<$ 1\si{\milli\meter}, 4\si{\milli\meter} $<L_{c}<$ 6\si{\milli\meter}, and 6\si{\milli\meter} $<L_{c}<$ 12\si{\milli\meter} particle sizes, respectively. The third observation is that ignition was not observed beyond 100s for either of the solid plate fuel bed materials. Trends in ignition times for the plates were most similar to those for beds with the smallest particles. Fourth, for the ignition events that occurred within the first 10\si{\second} there is no apparent relationship between time to ignition, temperature, particle size, and fuel bed type. Additionally, the long timescales of some ignition events suggest that smoldering initiates and then transitions to flaming combustion. Since the incidence of ignition at extended times increases as the particle size increases the potential for smoldering to flaming transition is attributed to thermal and physical properties of the fuel bed. The different sensitivities of ignition just described are attributed to differences in the bulk thermal properties, the interface between the heater and fuel bed, and the global equivalence ratio of the fuel bed, as explained later. 
        \begin{figure}[htp]
            \centering
            \begin{tikzpicture}
                \node[anchor=south west,inner sep=0] (image) at (0,0) {\includegraphics[width=\figureWidthSet]{Figures/ignition_comparison.png}};
                    \begin{scope}[x={(image.south east)},y={(image.north west)}]
                        \draw[dashed, thick,rounded corners] (0.12, 0.17) rectangle (0.33, 0.80);
                        \draw[dotted, thick,rounded corners] (0.46, 0.17) rectangle (0.68, 0.80);
                    \end{scope}
            \end{tikzpicture}
            % \includegraphics[width=\figureWidthSet]{Figures/ignition_comparison.png}
            \caption{Time to ignition and heater temperature at ignition for all fuel bed materials. 
            The dashed and dotted boxes emphasize the two general times-scales associated with ignition.}
            \label{fig:ignTime}
        \end{figure}

@@

\section{Summary and Conclusions}
    Flaming ignition tests have been conducted for porous Douglas-fir beds, Douglas-fir plates, and cardboard plates. A cylindrical cartridge heater was used as a firebrand surrogate. Heater temperature and electrical power to the heater were collected throughout each test. The derived heat flux to the fuel bed was within the range reported in literature of heat fluxes delivered by firebrands.  The time to ignition and probablity of ignition were used to evaluate the ignition propensity for the various fuel beds and heater temperatures.  A simplified heat transfer, pyrolysis, and ignition delay model was developed and used to provide further insights into the physical processes associated with ignition. The specific conclusions from this work are as follows.
        \begin{enumerate}
            \item 
                Smaller particles ignite more readily in porous beds than larger particles when heat transfer from the heater is primarily through conduction. This was evident by higher ignition probabilities, in general, of the smaller particles for a fixed heater temperature. As particle sizes increase radiant heat transfer becomes more important and fuel beds with larger particles were more likely than smaller particles to ignite at extended times (\textgreater 100\si{\second}) due to the increased importance of radiant ignition. 
            \item
                Douglas-fir plates ignite at times where conduction is the dominant mode of heat transfer (\textless 10\si{\second}) due to the higher thermal conductivity of the solid plates. The ignition probability of plates was the most similar to the larger particle, in particular at lower heater temperatures, due to dispersed heating of the porous fuel bed through radiation and the increased thermal conductivity of the plates creating similar temperature profiles. The rise in ignition probability  over a smaller heater temperature range time with temperature results from more consistent contact between the heater and plate surface.
            \item 
                Heat flux delivered to the fuel bed, when compared to heater temperature, is more indicative of ignition likelihood and ignition time for porous fuel beds. Heat flux is a more significant predictor of ignition because it captures differences in heat transfer modes and particle contact that heater temperature values do not. While this finding is not new, what is novel is that the mixed mode of heating (conduction and radiation) has a significant impact on the flaming ignition of fuel beds.
            \item 
                Consideration of the transport characteristics of pyrolyzate gases near the high temperature source can be important for more fully predicting ignition propensity. A \textit{Da} of ignition, in relation to the measured heat flux and thermal diffusivity of the fuel beds, is a promising relationship for predicting ignition for the porous fuel beds.  
        \end{enumerate}
    Further work is needed to verify that the \textit{Da} may be used to predict ignition for solid surfaces and for porous fuel beds with varying chemical compositions.  If proven valid, the (\textit{Da}), measured/predicted heat fluxes, and fuel bed properties may be used to help predict ignition of fuel beds both in and out of the WUI, ultimately helping to increase the effectiveness of fire prevention and suppression efforts.

### 2022-03-30 manuscript3.tex
Updates to manuscripts
Multi-heater introduction draft completed. It needs significant work but the structure is there.

\section{Abstract}
    The increasing severity of wildfires and the expansion of the wildland urban interface have placed a greater number of homes at risk for being destroyed by wildfires. Ignition of fuel beds on or near homes by embers is a significant source of fire spread and structure loss. Fire prevention and suppression efforts lack an expedient method for estimating the likelihood of ignition for the wide variety of materials around homes.

\section{Introduction}\addvspace{10pt}
\label{sec:introduction}
    Structure losses due to wildfires have increased markedly in recent years, with over 60\% of homes lost in the last 15 years occurring since 2017~\cite{Barrett2020}. One mechanism that leads to home loss during wildfires is ignition of flammable materials around a home (e.g., pine needles accumulated on a roof) by an ember~\cite{Syphard2019, Suzuki2020, Koo2010a}. The combustion of the flammable materials then ignite the structure itself, resulting the loss of the structure. When seeking to understand the processes that occur leading up to the ignition of a structure by firebrands three processes must be considered: (1) the generation of embers in the fire itself, (2) the transport of the embers to the fuel bed, and (3) the ignition of a fuel bed after the ember lands. This work focuses on event (3), specifically the processes of heat transfer and subsequent ignition that occur after an ember lands on a fuel bed.

@@

\section{Background}
\label{sec:background}
    Studies considering ignition have primarily focused on the influence of heat transfer and fuel bed conditions when identifying ignition parameters and the observed sensitivities to fuel bed properties are attributed to both heat transfer and chemistry related processes. The probability of ignition generally increases as the energy content of an ember increases~\cite{Hadden2011}. However, further research has indicated that energy content alone is not a sufficient indicator of ignition as a minimum ember temperature (a driver of heat flux rates) is also necessary~\cite{Zak2014}. Additionally, heat loss from embers to the surroundings has a significant effect on ignition~\cite{Fernandez-Pello2015}. While the aforementioned efforts have focused primarily on the impact of ember properties on ignition, efforts focused on the properties of the fuel bed have shown sensitivities to fuel bed material type and particle size~\cite{Urban2018}. Additionally, for fuel beds of constant particle size, an increase in the packing density of the fuel bed resulted in a decrease in the critical heat flux needed for ignition~\cite{Mindykowski2011, Hernandez2018, Rivera2020}. The observed sensitivities to fuel bed properties are likely caused by both heat transfer and chemistry related processes. Changes in the particle size of a fixed material may impact the contact area (e.g., a ball resting on fine sawdust compared to a pile of pine needles) between the ember and fuel bed; potentially reducing the amount of heat transferred through conduction as the contact area decreases. Additionally, when considering a fixed particle size fuel bed, changes in chemical composition may change the heat transfer and heating rate of the materials due to changes in the thermal conductivity of the material affecting the ignition properties of the material. For these reasons, knowledge of the effects of chemical composition on ignition, as it relates to thermal and chemical processes, must be addressed to accurately represent ignition properties for the wide range of materials found in fires resulting in home loss.

\section{Methods}
\label{sec:methods}
    The 50\% probability of flaming ignition with respect to the surface temperature of a resistance heater was determined for five different materials. The fuel bed materials used were Douglas-fir wood, pine wood, Douglas-fir bark, wheat straw, and oak wood. These materials were chosen to represent a range of chemical compositions, and are materials that may be subject to ember storms in a wildland urban interface fire. Table \ref{tab:composition} shows the estimated chemical composition of the materials based on the Bioenergy Feedstock Library~\cite{feedstock}. To create fuel beds of consistent particle size,  materials were granulated and then sorted such that the particles fit through a screen with openings of 2.1\si{\milli\meter} but not through openings of 0.85\si{\milli\meter}. Once sorted, the materials were oven dried at 103\si{\celsius}.
    \begin{table*}[hpbt]
        \caption{Proportion of cellulose, hemicellulose, and lignin of the materials tested estimated from the Bioenergy Feedstock Library~\cite{feedstock}}
        \centering
        \begin{tabular}{crrrr}
            % \rowcolor{gray!50}
            Material & Cellulose (\%) & Hemicellulose (\%) & Lignin (\%) & Other (\%) \\
            \hline
            Oak Wood         & 45 & 22 & 24 & 9\\
            Douglas-fir Wood & 44 & 22 & 28 & 6\\
            Wheat Straw      & 44 & 27 & 19 & 10\\
            Pine Wood        & 40 & 30 & 20 & 10\\
            Douglas-Fir Bark & 26 & 11 & 59 & 4
        \end{tabular}
        \label{tab:composition}
    \end{table*}

@@

\section{Results and Discussion}
    \label{sec:results}
    The flaming or non-flaming ignition outcome of each test and the logistic regression results for Douglas-fir wood, oak wood, and pine wood are shown in Figure~\ref{fig:logistic_plot}. The circular markers represent the outcome of each test, either flaming or no ignition. The logistic regression for each material is shown as a solid line with the shaded regions representing the 95\% confidence interval for each regression. Douglas-fir bark and wheat straw are not shown in Figure~\ref{fig:logistic_plot} because flaming ignition was not observed for five tests at the maximum heater temperature of 750\si{\celsius}. The heater temperature estimated to produce 50\% ignition probability from the logistic regressions is 574\si{\celsius} for the Douglas-fir wood, 634\si{\celsius} for the pine wood, and 662\si{\celsius} for the oak wood. Anecdotally, self sustained smoldering was observed for all five materials at temperatures lower than the flaming ignition temperature. The heater temperature corresponding the onset of smoldering ignition was not measured in this study. 
        \begin{figure}[htpb]
            \centering
            \includegraphics[width=\figureWidthSet, trim={0 0 37cm 0}, clip]{conference_results_binary.png}
            \caption{Test results for materials where flaming ignition was observed. The circular markers denote individual tests. The solid lines represent the logistic regression and the shaded zones the 95\% confidence interval with the 50\% probability of ignition labelled}
            \label{fig:logistic_plot}
        \end{figure}
    Two aspects of the ignition results are of note. As anticipated, the differences in estimated heater temperature to produce 50\% ignition probability suggest that significant differences in ignition occur across a range of materials in the same apparatus and experimental conditions. Second, the transition between no ignition and flaming ignition may become less defined (i.e., the 95\% confidence interval of the regression spans a much larger temperature range) as the ignition temperature decreases. This is significant because it may cause more difficulty in predicting the ignition of materials that ignite at lower temperatures. Lower confidence in predicting ignition for easily ignitable materials would be detrimental to the usefulness of a model or predictive tool that may be implemented from this testing methodology. It is, however, unclear if the ignition to no ignition transition in other materials of similarly low ignition temperature will behave similarly to that of Douglas-fir. Characterization of additional materials is needed to elucidate these trends.

>>>>

\section{Abstract}
\label{sec:abtract3}
    The increasing severity of wildfires and the expansion of the wildland urban interface have placed a greater number of homes at risk for being destroyed by wildfires. Ignition of fuel beds on or near homes by embers is a significant source of fire spread and structure loss. Fire prevention and suppression efforts lack an expedient method for estimating the likelihood of ignition for the wide variety of materials around homes.

\section{Introduction}\addvspace{10pt}
\label{sec:introduction3}
    Structure losses due to wildfires have increased markedly in recent years, with over 60\% of homes lost in the last 15 years occurring since 2017~\cite{Barrett2020}. One mechanism that leads to home loss during wildfires is ignition of flammable materials around a home (e.g., pine needles accumulated on a roof) by an ember~\cite{Syphard2019, Suzuki2020, Koo2010a}. The combustion of the flammable materials then ignite the structure itself, resulting the loss of the structure. When seeking to understand the processes that occur leading up to the ignition of a structure by firebrands three processes must be considered: (1) the generation of embers in the fire itself, (2) the transport of the embers to the fuel bed, and (3) the ignition of a fuel bed after the ember lands. This work focuses on event (3), specifically the processes of heat transfer and subsequent ignition that occur after an ember lands on a fuel bed.

@@

\section{Background}
\label{sec:background3}
    Studies considering ignition have primarily focused on the influence of heat transfer and fuel bed conditions when identifying ignition parameters and the observed sensitivities to fuel bed properties are attributed to both heat transfer and chemistry related processes. The probability of ignition generally increases as the energy content of an ember increases~\cite{Hadden2011}. However, further research has indicated that energy content alone is not a sufficient indicator of ignition as a minimum ember temperature (a driver of heat flux rates) is also necessary~\cite{Zak2014}. Additionally, heat loss from embers to the surroundings has a significant effect on ignition~\cite{Fernandez-Pello2015}. While the aforementioned efforts have focused primarily on the impact of ember properties on ignition, efforts focused on the properties of the fuel bed have shown sensitivities to fuel bed material type and particle size~\cite{Urban2018}. Additionally, for fuel beds of constant particle size, an increase in the packing density of the fuel bed resulted in a decrease in the critical heat flux needed for ignition~\cite{Mindykowski2011, Hernandez2018, Rivera2020}. The observed sensitivities to fuel bed properties are likely caused by both heat transfer and chemistry related processes. Changes in the particle size of a fixed material may impact the contact area (e.g., a ball resting on fine sawdust compared to a pile of pine needles) between the ember and fuel bed; potentially reducing the amount of heat transferred through conduction as the contact area decreases. Additionally, when considering a fixed particle size fuel bed, changes in chemical composition may change the heat transfer and heating rate of the materials due to changes in the thermal conductivity of the material affecting the ignition properties of the material. For these reasons, knowledge of the effects of chemical composition on ignition, as it relates to thermal and chemical processes, must be addressed to accurately represent ignition properties for the wide range of materials found in fires resulting in home loss.

\section{Methods}
\label{sec:methods3}
    The 50\% probability of flaming ignition with respect to the surface temperature of a resistance heater was determined for five different materials. The fuel bed materials used were Douglas-fir wood, pine wood, Douglas-fir bark, wheat straw, and oak wood. These materials were chosen to represent a range of chemical compositions, and are materials that may be subject to ember storms in a wildland urban interface fire. Table \ref{tab:composition} shows the estimated chemical composition of the materials based on the Bioenergy Feedstock Library~\cite{feedstock}. To create fuel beds of consistent particle size,  materials were granulated and then sorted such that the particles fit through a screen with openings of 2.1\si{\milli\meter} but not through openings of 0.85\si{\milli\meter}. Once sorted, the materials were oven dried at 103\si{\celsius}.
    \begin{table*}[hpbt]
        \caption{Proportion of cellulose, hemicellulose, and lignin of the materials tested estimated from the Bioenergy Feedstock Library~\cite{feedstock}}
        \centering
        \begin{tabular}{crrrr}
            % \rowcolor{gray!50}
            Material & Cellulose (\%) & Hemicellulose (\%) & Lignin (\%) & Other (\%) \\
            \hline
            Oak Wood         & 45 & 22 & 24 & 9\\
            Douglas-fir Wood & 44 & 22 & 28 & 6\\
            Wheat Straw      & 44 & 27 & 19 & 10\\
            Pine Wood        & 40 & 30 & 20 & 10\\
            Douglas-Fir Bark & 26 & 11 & 59 & 4
        \end{tabular}
        \label{tab:composition}
    \end{table*}

@@

\section{Results and Discussion}
    \label{sec:results3}
    The flaming or non-flaming ignition outcome of each test and the logistic regression results for Douglas-fir wood, oak wood, and pine wood are shown in Figure~\ref{fig:logistic_plot}. The circular markers represent the outcome of each test, either flaming or no ignition. The logistic regression for each material is shown as a solid line with the shaded regions representing the 95\% confidence interval for each regression. Douglas-fir bark and wheat straw are not shown in Figure~\ref{fig:logistic_plot} because flaming ignition was not observed for five tests at the maximum heater temperature of 750\si{\celsius}. The heater temperature estimated to produce 50\% ignition probability from the logistic regressions is 574\si{\celsius} for the Douglas-fir wood, 634\si{\celsius} for the pine wood, and 662\si{\celsius} for the oak wood. Anecdotally, self sustained smoldering was observed for all five materials at temperatures lower than the flaming ignition temperature. The heater temperature corresponding the onset of smoldering ignition was not measured in this study. 
        \begin{figure}[htpb]
            \centering
            \includegraphics[width=\figureWidthSet, trim={0 0 37cm 0}, clip]{conference_results_binary.png}
            \caption{Test results for materials where flaming ignition was observed. The circular markers denote individual tests. The solid lines represent the logistic regression and the shaded zones the 95\% confidence interval with the 50\% probability of ignition labelled}
            \label{fig:logistic_plot}
        \end{figure}
    Two aspects of the ignition results are of note. As anticipated, the differences in estimated heater temperature to produce 50\% ignition probability suggest that significant differences in ignition occur across a range of materials in the same apparatus and experimental conditions. Second, the transition between no ignition and flaming ignition may become less defined (i.e., the 95\% confidence interval of the regression spans a much larger temperature range) as the ignition temperature decreases. This is significant because it may cause more difficulty in predicting the ignition of materials that ignite at lower temperatures. Lower confidence in predicting ignition for easily ignitable materials would be detrimental to the usefulness of a model or predictive tool that may be implemented from this testing methodology. It is, however, unclear if the ignition to no ignition transition in other materials of similarly low ignition temperature will behave similarly to that of Douglas-fir. Characterization of additional materials is needed to elucidate these trends.
